---
title: "HW06_Diffusion"
author: "Dani Sclafani, Julia Parish, Kristin Gill"
date: '2022-05-05'
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
output: 
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


### Load Libraries

```{r}
library(tidyverse)
library(deSolve)
library(here)
library(cowplot)
```


# Source Diffusion Function

```{r}
source(here("R/diffusion.R"))
```

Diffusion Model Parameters

- `initialC` = initial concentration (mg/L)
- `dx` = length of each segment (m)
- `nx` = number of discrete segments (m)
- `nt` = number of discrete time intervals (s)
- `dt` =  seconds in each time interval (s)
- `area` = area of cross section of container (m2)
- `D` = diffusivity (how easily the chemical diffuses (s/m2)  

# Run the in class diffussion model examples
```{r}
# run our diffusion model (iterative difference equation) with initial concentration of 10, for 8 timestep (size 1m), and 10 space steps (size 1s)

# using diffusion parameters 0.5 s/m2, 10 m2
result = diff1(initialC = 10, nx = 10, dx = 1, nt = 8, dt = 1, D = 0.5, area = 10)

# a list is returned with our 3 data frames for concentration (conc), qin and qout
result
```

```{r}
# used filled contour to plot results
head(result$conc)

filled.contour(result$conc, xlab="Time", ylab="Distance")

# or if you prefer this orientation (Distance on x axis)
filled.contour(t(result$conc), ylab="Time", xlab="Distance")
```

# In class example of changing parameters (diffusivity D, and space and time steps (dx, dt))

```{r}
# changes diffusivity and other parameters particularly
# diffusivity, dx and dt

res = diff1(initialC = 10, nx = 10, dx = 1, nt = 10, dt = 30, D = 0.006, area = 1)

filled.contour(res$conc, xlab="Time", ylab="Distance")

# we can also see how much material moved from place to place each time step
filled.contour(res$qin, xlab="Time", ylab="Distance")

```

# Homework Responses

# Change diffusivity and length of segment

```{r}
# changes diffusivity (D) and length of segment (dx)
# original params: initialC = 10, nx = 10, dx = 1, nt = 8, dt = 1, D = 0.5, area = 10
# increased segment length to 4 m, diffusivity to 1 s/m2

resultA = diff1(initialC = 10, nx = 10, dx = 4, nt = 8, dt = 1, D = 1, area = 10)

resultA_plot1 <- filled.contour(resultA$conc, 
               color.palette = function(n) hcl.colors(n, "GnBu", rev = TRUE),
               plot.title = title(main = "Diffusion - Change in Diffusivity and Segment Length", xlab="Time", ylab = "Distance"))

resultA_plot1
```

```{r}
# how much material moved from place to place each time step
resultA_plot2 <-filled.contour(resultA$qin, 
               color.palette = function(n) hcl.colors(n, "GnBu", rev = TRUE),
               plot.title = title(main = "Diffusion Movement - Change in Diffusivity & Segment Length", xlab="Time", ylab="Distance"))

resultA_plot2
```


# Change space step (dx), time step (dt)

```{r}
# changes length of segment (dx) and time (dt)
# original params: initialC = 10, nx = 10, dx = 1, nt = 8, dt = 1, D = 0.5, area = 10
# decreased length of segment to 0.5m, and time in each interval to 0.5 second

resultb = diff1(initialC = 10, nx = 10, dx = .5, nt = 8, dt = .5, D = 0.5, area = 10)

resultB_plot1 <- filled.contour(resultb$conc, 
               color.palette = function(n) hcl.colors(n, "BuPu", rev = TRUE),
               plot.title = title(main = "Diffusion - Change in Time & Segment Length", xlab="Time", ylab = "Distance"))

resultB_plot1
```


```{r}
# how much material moved from place to place each time step
resultB_plot2 <-filled.contour(resultb$qin, 
               color.palette = function(n) hcl.colors(n, "BuPu", rev = TRUE),
               plot.title = title(main = "Diffusion Movement - Change in Time & Segment Length", xlab="Time", ylab="Distance"))

resultB_plot2
```

- `initialC` = initial concentration (mg/L)
- `dx` = length of each segment (m)
- `nx` = number of discrete segments (m)
- `nt` = number of discrete time intervals (s)
- `dt` =  seconds in each time interval (s)
- `area` = area of cross section of container (m2)
- `D` = diffusivity (how easily the chemical diffuses (s/m2)  


# Change all parameters
 
```{r}
# original params: initialC = 10, nx = 10, dx = 1, nt = 8, dt = 1, D = 0.5, area = 10
# increased initial concentration to 50 mg/L, number of segments from 10 to 30, length of segment (dx) to from 1 m to 6 m, # of time intervals to 10 from 8, seconds in each time interval from 1 sec to 10 sec, diffusivity to 0.7 s/m2 and the area of container from 10 m2 to 60 m2.

# making slight changes to dx (length of segment), dt (secs in each time interval) had significant influences to the diffusion rates in this model. 

resultc = diff1(initialC = 50, nx = 30, dx = 6, nt = 10, dt = 10, D = .7, area = 60)

resultC_plot1 <- filled.contour(resultc$conc, 
               color.palette = function(n) hcl.colors(n, "spectral", rev = TRUE),
               plot.title = title(main = "Diffusion - Change in All Parameters", xlab="Time", ylab = "Distance"))

resultC_plot1
```

```{r}
# how much material moved from place to place each time step
resultC_plot2 <-filled.contour(resultc$qin, 
               color.palette = function(n) hcl.colors(n, "spectral", rev = TRUE),
               plot.title = title(main = "Diffusion Movement - Change in All Parameters", xlab="Time", ylab="Distance"))

resultC_plot2
```



